The SMT Competition 2015--2018

### 2018-11-20 2018/report/SMT-COMP.tex
medals
Introduction

@@

Satisfiability Modulo Theories (SMT) is a generalization of Boolean
satisfiability (SAT), the satisfiability decision problem for
propositional logic.  In place of Boolean variables, SMT formulas may
contain terms that are built from function and predicate symbols drawn
from a number of background theories.  Background theories, which are
motivated by application domains, include the theory of
arrays~\cite{TODO}, integer and real arithmetic~\cite{TODO},
bit-vectors~\cite{TODO}, and floating-point numbers~\cite{TODO} among
others.  For instance, the following is an SMT formula over the
combination of integer arithmetic and uninterpreted functions:
%
$$x \leq y \wedge y \leq x \wedge P (f(x) - f(y)) \wedge \neg P(0)$$
%
Software tools to determine the satisfiability of such formulas are
called SMT solvers.  With its rich input language, SMT has
applications in software engineering, optimization, and many other
areas~\cite{DeMoura:2011:SMT}.

Internally, many SMT solvers employ SAT solving techniques to deal
with the propositional structure of the formula, and combine these
with (semi-)decision procedures for the background theories that the
solver supports~\cite{TODO,TODO,TODO}.  Historically, SMT solvers have
focused on quantifier-free formulas and decidable combinations of
background theories.  Increasingly, however, SMT solvers also support
quantified formulas.  Hence there is overlap between SMT solving and
automated theorem proving for quantified Boolean formulas (QBF),
first-order logic, and even higher-order logic~\cite{TODO}.

The International Satisfiability Modulo Theories Competition
(SMT-COMP) is an annual competition between SMT solvers.  It was
instituted in~2005~\cite{BdMS05}, and is affiliated with the
International Workshop on Satisfiability Modulo Theories.  Solvers are
submitted to the competition by their developers, and pitted against
each other in a number of tracks and divisions.  The SMT Competition
was part of the Federated Logic Conference (FLoC) Olympic Games, which
comprised 14 competitions in various areas of computational logic, in
Vienna in~2014, and again in Oxford~(UK) in~2018.  It was last
described in a 2014 competition
report~\cite{DBLP:journals/jsat/CokDW14}.  In this paper, we report on
the design and selected results of the competition during the last
FLoC Olympiad, from 2015 to~2018.

\begin{tjark}
  WIP

  structure of the paper
\end{tjark}

@@

SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers.''~\cite{DBLP:journals/jar/CokSW15} Together with the
increased computational resources available to the competition since
its move to StarExec, this prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible
benchmarks. \TODO: benchmark weights

@@

Since 2014 (and previously for the evaluation in
2013~\cite{DBLP:journals/jar/CokSW15}), the computational workload of
the competition is being run on StarExec, a ``cross-community
infrastructure for logic solving''~\cite{DBLP:conf/cade/StumpST14}
developed at the University of Iowa and supported by the National
Science Foundation.  Each machine in the StarExec cluster is equipped
with two E5-2609 Intel Xeon processors (10~MiB Cache, 2.4~GHz).  This
hardware configuration has been unchanged since~2013.  Up to two
solvers are run simultaneously on each machine: one solver on each
processor, with four cores available.  Memory usage is limited
to~60~GiB per solver.

@@

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver-benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~72\,\%, from~1,028,615 job pairs in~2015 to~1,776,062 job
pairs in~2018.  (This is the number of unique solver-benchmark
combinations that were part of the competition.  The number of job
pairs that were actually run is even higher: test jobs and job pairs
that had to be re-run, e.g., because of issues with the StarExec
infrastructure, are not included.)  For comparison, the~2014
competition only ran~339,714 job
pairs~\cite{DBLP:journals/jsat/CokDW14}.

@@

By number of solver versions submitted, the competitions in 2015--2018
were the four largest in the history of SMT-COMP.  On average, 26
solver versions were submitted each year since 2015.  In contrast, the
competitions from 2005--2014 only received an average of 12
submissions~\cite{DBLP:journals/jsat/CokDW14}.

@@

The 2018 edition of the competition was again part of the FLoC Olympic
Games, which sponsored seven FLoC 2018 medals.  The competition
organizers awarded these to the following solvers:
\begin{itemize}
\item The best three main-track solvers according to the
  competition-wide ranking (first place: CVC4, second place:
  Yices-2.6.0, third place: SMTInterpol);
\item the solver(s) that won the QF\_BV division in the main track
  (sequential ranking: Boolector, parallel ranking: Minkeyrink-MT);
\item the solver that won the most competitive divisions in the
  application track (Yices-2.6.0);
\item the solver that won the most competitive divisions in the
  unsat-core track (Yices-2.6.0).
\end{itemize}

>>>>

%\usepackage[numbers,sort&compress]{natbib}

@@

Satisfiability Modulo Theories (SMT) is a generalization of Boolean
satisfiability (SAT), the satisfiability decision problem for
propositional logic.  In place of Boolean variables, SMT formulas may
contain terms that are built from function and predicate symbols drawn
from a number of background theories.  Background theories, which are
motivated by application domains, include the theory of arrays,
integer and real arithmetic, bit-vectors, and floating-point numbers
among others~\cite{BarFT-RR-17}.  For instance, the following is an
SMT formula over the combination of integer arithmetic and
uninterpreted functions:
%
$$x \leq y \wedge y \leq x \wedge P (f(x) - f(y)) \wedge \neg P(0)$$
%
Software tools to determine the satisfiability of such formulas are
called SMT solvers.  With its rich input language, SMT has
applications in software engineering, optimization, and many other
areas~\cite{DeMoura:2011:SMT}.

Internally, many SMT solvers employ SAT solving techniques to deal
with the propositional structure of the formula, and combine these
with (semi-)decision procedures for the background theories that the
solver
supports~\cite{DBLP:conf/cav/BarrettCDHJKRT11,DBLP:conf/cade/BoutonODF09,DBLP:conf/tacas/MouraB08}.
Historically, SMT solvers have focused on quantifier-free formulas and
decidable combinations of background theories.  Increasingly, however,
SMT solvers also support quantified formulas.  Hence there is overlap
between SMT solving and automated theorem proving for quantified
Boolean formulas (QBF), first-order logic, and even higher-order
logic~\cite{DBLP:journals/corr/abs-1712-01486}.

The International Satisfiability Modulo Theories Competition
(SMT-COMP)~\cite{BdMS05,BdMS07,BDdMOS13,BDOS08,BDOS10,CDW14,CGBD12} is
an annual competition between SMT solvers.  It was instituted in~2005,
and is affiliated with the International Workshop on Satisfiability
Modulo Theories.  Solvers are submitted to the competition by their
developers, and pitted against each other in a number of tracks and
divisions.  The SMT Competition was part of the Federated Logic
Conference (FLoC) Olympic Games, which comprised 14 competitions in
various areas of computational logic, in Vienna in~2014, and again in
Oxford~(UK) in~2018.  It was last described in a 2014 competition
report~\cite{CDW14}.  In this paper, we report on the design and
selected results of the competition during the last FLoC Olympiad,
from 2015 to~2018.

For space reasons, the discussion in this paper is necessarily
limited.  Complete data about past competitions since 2015 (and partly
for earlier competitions) is available from the competition web site,
\url{http://www.smtcomp.org}.  This includes detailed results for each
solver/benchmark combination, totaling nearly 6 million data records
(about 1.1~GB in CSV format).

The rest of this paper is structured as follows.  In
Section~\ref{sec:goals}, we present the goals and organization of the
competition.  Section~\ref{sec:smtlib} introduces the SMT-LIB language
and library, and Section~\ref{sec:benchmarks} discusses competition
divisions and benchmarks.  The competition is being run on StarExec,
which is described in Section~\ref{sec:starexec}.
Section~\ref{sec:participants} gives an overview of participating
solvers.  The details of the competition procedure are described in
Section~\ref{sec:procedure}, and selected results are presented and
analyzed in Section~\ref{sec:results}.  We conclude with suggestions
for future SMT Competitions in Section~\ref{sec:conclusions}.

@@

SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers.''~\cite{CSW15} rTogether with the increased
computational resources available to the competition since its move to
StarExec, this prompted a paradigm shift.  Since 2015, the competition
is evaluating all solvers on all eligible benchmarks. \TODO: benchmark
weights

@@

Since 2014 (and previously for the evaluation in 2013~\cite{CSW15}),
the computational workload of the competition is being run on
StarExec, a ``cross-community infrastructure for logic
solving''~\cite{DBLP:conf/cade/StumpST14} developed at the University
of Iowa and supported by the National Science Foundation.  Each
machine in the StarExec cluster is equipped with two E5-2609 Intel
Xeon processors (10~MiB Cache, 2.4~GHz).  This hardware configuration
has been unchanged since~2013.  Up to two solvers are run
simultaneously on each machine: one solver on each processor, with
four cores available.  Memory usage is limited to~60~GiB per solver.

@@

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver-benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~72\,\%, from~1,028,615 job pairs in~2015 to~1,776,062 job
pairs in~2018.  (This is the number of unique solver-benchmark
combinations that were part of the competition.  The number of job
pairs that were actually run is even higher: test jobs and job pairs
that had to be re-run, e.g., because of issues with the StarExec
infrastructure, are not included.)  For comparison, the~2014
competition only ran~339,714 job pairs~\cite{CDW14}.

@@

By number of solver versions submitted, the competitions in 2015--2018
were the four largest in the history of SMT-COMP.  On average, 26
solver versions were submitted each year since 2015.  In contrast, the
competitions from 2005--2014 only received an average of 12
submissions~\cite{CDW14}.

@@

The 2018 edition of the competition was again part of the FLoC Olympic
Games, which sponsored seven medals.  The competition organizers
awarded these to the following solvers:
\begin{itemize}
\item The best three main-track solvers according to the
  competition-wide ranking (first place: CVC4, second place:
  Yices-2.6.0, third place: SMTInterpol);
\item the solver(s) that won the QF\_BV division in the main track
  (sequential ranking: Boolector, parallel ranking: Minkeyrink-MT);
\item the solver that won the most competitive divisions in the
  application track (Yices-2.6.0);
\item the solver that won the most competitive divisions in the
  unsat-core track (Yices-2.6.0).
\end{itemize}
